Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2024-10-10 main.tex
some progress... need to get some sleep now ;)

    %%% MOTIVATION %%%
    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
    and memory sizes. In general, having different memory sizes per CPUs makes it more challenging to compute
    a schedule that respects all memory constraints -- meaning that no task is executed on a 
    processor with less memory than needed for the task. This is, however, very important to
    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
    Hence, building on previous related %\hmey{You may want to add other refs not from us} 
    work~\cite{gou2020partitioning,He21,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
    formulation that takes memory sizes as explicit constraints into account. Its objective is
    the very common \emph{makespan}~\cite{liu2018survey}, 
    which acts as proxy for the total execution time of a workflow.

@@

    Another limitation in practice of HEFT (and many other scheduling strategies) is their 
    assumption that the task running times provided to them are accurate. In practice, this is 
    not the case and deviations from user estimates or historical measurements are 
    very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
    deviations occur. List-based schedulers such as HEFT are, however, not designed for 
    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
%    and would compute a completely new schedule from scratch.

    %%% CONTRIBUTION %%%
    \paragraph*{Contribution} In this paper, we first propose three HEFT-based heuristics
    that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
    (M behind HEFT for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
    BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
    and MM for \underline{m}inimum \underline{m}emory traversal).
    The difference between the new heuristics is the way they prioritize tasks (for processor assignment).
    We compare the static schedule produced by these heuristics with the classical HEFT as
    baseline, which however does not take memory sizes into account. 
    \hmey{Need to clarify: why is this SotA?}

    \begin{itemize}
      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
    \end{itemize}

@@

    So, during the task prioritization phase, Ref.~\cite{sulaiman2021hybrid} computes the standard deviation of the computation cost
    (between processors) and adds it to the mean value to account for the differences between processor speeds.
    In the processor choice phase, they duplicate the entry task and the longest parent tasks during idle times on the processor.

    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
    the fastest and the slowest processors, divided by the speed ratio of these two processors.
    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
    They choose the processor with the lowest execution time and cross over to other processors sometimes.
    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}

@@

    The TSHCS (Task Scheduling for Heterogeneous Comupting Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
    by adding randomized decisions to the second phase.
    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
    produces the lowest finish time.

@@

    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement
    learning techniques(\cite{yano2022cqga}).

    \subsection{Memory-aware scheduling algorithms}
    Respecting processor memories adds a constraint to a scheduling problem.
    Therefore, only specifically memory-targeted algorithms address this issue.
    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
    is formulated and addressed in the algorithm.

    Different models of memory available on processors and memory requirements of tasks have been presented.

@@

    In an assumed dual-memory systems(\cite{herrmann2014memory}) a processor can have access to a memory of two different
    kinds (red or blue), and each task can be executed on only one sort of memory.
    The communications happen only between these two kinds of processors (communications withing each group are ignored).
    The authors then formulate an ILP solution for this problem formulation.

    The algorithm presented by Yao \etal~\cite{yao2022memory}, each processor has an own internal memory and all
    processors share a common external one. The internal (local) memory is used to store the task files.
    The external memory is used to store evicted files to make room for the execution of a task on a processor.
    All processors, including the original one, can access these files.
    Each edge in~\cite{yao2022memory} has two weights -- the size of the files transferred along it,
    and the time of communication along this edge.
    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.

    \cite{ding2024ils} assumes connected processors with individual limited memories.
    The collective set of memories forms the global memory, to which each processor has access, however the access time
    to global memory is different.
    Each memory access in the graph is modelled as a memory access token on the task, while the edges have no weights.
    The solved problem is how to allocate initial input data in processor memories so that the overall
    execution is minimized and the memories are not exceeded.
    They propose an integer linear programming model.
    %that minimizes the length of the critical path, including a greedy initial solution.

@@

    Cloud-oriented model can include costs associated with memory usage(\cite{liang2020memory}).

@@

    DVR HEFT~\cite{SANDOKJI2019482} uses an almost unchanged HEFT algorithm in their static step, executing three slightly
    varying variants of task weighting and choosing the variant that gives the best overall makespan.
    In the dynamic phase, they receive new tasks and schedule them on either idle processors or those processors that give them
    the earliest finish time.
    Task failures are not covered.

@@

    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
    on a single machine.
    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
    that these runtime changes are unrelated to each other).
    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
    performance measure $\phi$.
    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.

    \subsection{Other notable works}

    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
    They utilize task duplication when creating the clusters (grains).
    Their scheduler then maps clusters to processors.
    They assume unlimited processors with the speed of 1.
    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
    to it as possible.
    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
    %release times.

    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
    to search for an optimal solution for mapping tasks on machines.
    A solution is generated at each step, and the best solution is kept at the end.
    The search terminates when a certain termination criterion is reached.
    It generates better results than other algorithms, because it explores the whole solution space.

    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
    The heuristic first determines similar activities and groups them together.
    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
    The heuristic calculates a new execution path with these tasks.

    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
    (processor) speed heterogeneity.
    Each accelerator has its own memory that can be zero.
    Each accelerator's characteristics depend on the task it runs and are not fixed.

    \cite{meng2018traffic} investigate scheduling on multi-core chips.
    Their model is far from ours.

    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
    They prioritize tasks that have failed before or are well-predictable.

    \section{Model} %: \skug{Full: 4, polished: 3}}

@@

    \section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
\label{sec.heuristics}

@@

    \section{Dynamic scenario}

@@

    \section{Experimental evaluation: \skug{Full: 3, Polished: 3}}
    \subsection{Setup: \skug{Full: 5, Polished: 4}}
    \label{sec:setup}

@@

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules \skug{THIS FIGURE IS WRONG}}
        \label{fig:mem-usages-onlyvalid}
        \vspace{-0.3cm}
    \end{figure}

@@

    \section{Conclusion  \skug{Full: 0, Polished: 0}}

>>>>

    %%% MOTIVATION %%%
    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
    and memory sizes. In general, having different memory sizes per CPUs makes it more challenging to compute
    a schedule that respects all memory constraints -- meaning that no task is executed on a 
    processor with less memory than needed for the task. This is, however, very important to
    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
    Hence, building on previous related %\hmey{You may want to add other refs not from us} 
    work~\cite{gou2020partitioning,He21,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
    formulation that takes memory sizes as explicit constraints into account. Its objective is
    the very common \emph{makespan}~\cite{liu2018survey}, 
    which acts as proxy for the total execution time of a workflow.
However, to the best of our knowledge,
    the only memory-aware heuristics that would account for memory constraints are  partitioning
    the DAG and not reusing processors once they have processed a part of the graph, leading
    to high values of makespan compared to a finer grain solution that reuses processors.

@@

    Another limitation in practice of HEFT (and many other scheduling strategies) is their 
    assumption that the task running times provided to them are accurate. In practice, this is 
    not the case and deviations from user estimates or historical measurements are 
    very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
    deviations occur. However, the original list-based schedulers, such as HEFT, are only defined
    in a static setting with accurate task parameters. 
%    
%    List-based schedulers such as HEFT are, however, not designed for 
%    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
%    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
%%    and would compute a completely new schedule from scratch.

    %%% CONTRIBUTION %%%
    %\paragraph*{Contribution} 
    The main contributions of this paper are both algorithmic and experimental:
    \begin{itemize}
\item We formalize the problem with memory constraints, where communication buffers
are used to evict data from memory if it will be later used by another processor. 
\item We design three HEFT-based heuristics
    that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
    (M behind HEFT for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
    BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
    and MM for \underline{m}inimum \underline{m}emory traversal).
    The difference between the new heuristics is the way they prioritize tasks (for processor assignment).

    \item We implement a runtime system able to provide some feedback to the scheduler
    when task requirements (in terms of execution time and memory) differ from the initial predictions, 
    and we recompute a schedule, based on the reported deviations.

    \item We perform extensive simulations, first in the static case by comparing the schedules produced 
    by these heuristics with the classical HEFT as baseline, which however does not take memory sizes into account; 
    while HEFT returns invalid schedules that exceed the processor memories and cannot execute correctly,
    the new heuristics are able to successfully schedule large workflows, with reasonable makespans.

    \item In the dynamic setting, we use a runtime system that allows us to simulate workflow executions,
    introducing deviations in running times and memory requirements of tasks that are communicated
    back to the scheduler; the scheduler can then recompute a schedule. Without these recomputations,
    most schedules become invalid after deviations, since the memory constraint is exceeded 
    for most workflows, hence demonstrating the necessity of a dynamic adjustment of the schedule.

%    \hmey{Need to clarify: why is this SotA?}

  %  \begin{itemize}
%      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
%      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
%      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
    \end{itemize}

We first review related work in Section~\ref{sec:related-work}. Then, we formalize the model in Section~\ref{sec:model}
and the algorithms in Section~\ref{sec:heuristics}. The adaptation of the heuristics in a dynamic setting is discussed in Section~\ref{sec:dyn}, and experiment results are presented in Section~\ref{sec:expe}. Finally, we conclude
and provide future working directions in Section~\ref{sec:conc}.

@@

    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
    In the processor choice phase, the entry task and the longest parent tasks are duplicated 
    during idle times on the processor.

%    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
%    the fastest and the slowest processors, divided by the speed ratio of these two processors.
%    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
%    They choose the processor with the lowest execution time and cross over to other processors sometimes.
%    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}

@@

    The TSHCS (Task Scheduling for Heterogeneous Computing Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
    by adding randomized decisions to the second phase.
    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
    produces the lowest finish time.

@@

    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement learning techniques~\cite{yano2022cqga}.

    \subsection{Memory-aware scheduling algorithms}
    Respecting processor memories adds a constraint to a scheduling problem.
    Therefore, only specifically memory-targeted algorithms address this issue.
    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
    is formulated and addressed in the algorithm.
%
    Different models of memory available on processors and memory requirements of tasks have been presented.

@@

    In an assumed dual-memory systems~\cite{herrmann2014memory},  a processor can have access 
    to a memory of two different
    kinds (red or blue), and each task can be executed on only one sort of memory.
    The communications happen only between these two kinds of processors (communications within 
    each group are ignored).
    The authors then formulate an ILP solution for this problem formulation.

    %The algorithm presented by Yao et al.
    Yao et al.~\cite{yao2022memory} consider that each processor has an own internal memory and all
    processors share a common external one. The internal (local) memory is used to store the task files.
    The external memory is used to store evicted files to make room for the execution of a task on a processor.
    All processors, including the original one, can access these files. 
    Each edge %in~\cite{yao2022memory} 
    has two weights -- the size of the files transferred along it,
    and the time of communication along this edge.
    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.

    In~\cite{ding2024ils}, there are connected processors with individual limited memories.
    The collective set of memories forms the global memory, to which each processor has access, however the access time
    to global memory is different.
    Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
    The solved problem is how to allocate initial input data in processor memories so that the overall
    execution is minimized and the memories are not exceeded.
    The authors propose an integer linear programming model.
    %that minimizes the length of the critical path, including a greedy initial solution.

@@

   Finally, there are some cloud-oriented models that include costs associated with memory usage~\cite{liang2020memory}.

   Overall, there are a variety of memory models, but, to the best of our knowledge, the only study on a multiprocessor
   platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24},
   but where a partition of the workflow is proposed, hence preventing processor reuse. 
   Hence, in ~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need of communication buffers to store data 
   that should be communicated
   between two processors when tasks are ready to execute.

@@

We now review related work in a dynamic setting. With no variation in task parameters, 
    DVR HEFT~\cite{SANDOKJI2019482} rather considers that new tasks arrive in the system. 
    They use an almost unchanged HEFT algorithm in the static step, executing three slightly
    varying variants of task weighting and choosing the variant that gives the best overall makespan.
    In the dynamic phase, they receive new tasks and schedule them on either idle processors or 
    those processors that give them
    the earliest finish time.
    %Task failures are not covered.

@@

    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
    on a single machine.
    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
    that these runtime changes are unrelated to each other).
    The authors formulate a decision space of all permutations of $n$ jobs, and the optimal schedule in relation to a
    performance measure $\phi$.
    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.

\AB{Say why our approach in dynamic setting is different and novel}

%    \subsection{Other notable works}
%
%    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
%    They utilize task duplication when creating the clusters (grains).
%    Their scheduler then maps clusters to processors.
%    They assume unlimited processors with the speed of 1.
%    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
%    to it as possible.
%    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
%    %release times.
%
%    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
%    to search for an optimal solution for mapping tasks on machines.
%    A solution is generated at each step, and the best solution is kept at the end.
%    The search terminates when a certain termination criterion is reached.
%    It generates better results than other algorithms, because it explores the whole solution space.
%
%    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
%    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
%    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
%    The heuristic first determines similar activities and groups them together.
%    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
%    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
%    The heuristic calculates a new execution path with these tasks.
%
%
%    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
%    (processor) speed heterogeneity.
%    Each accelerator has its own memory that can be zero.
%    Each accelerator's characteristics depend on the task it runs and are not fixed.
%
%
%
%
%    \cite{meng2018traffic} investigate scheduling on multi-core chips.
%    Their model is far from ours.
%
%
%    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
%    They prioritize tasks that have failed before or are well-predictable.
%
%

\section{Model} %: \skug{Full: 4, polished: 3}}
\label{sec:model}

@@

\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
\label{sec:heuristics}

@@

\section{Dynamic scenario}
\label{sec:dyn}

@@

\section{Experimental evaluation: \skug{Full: 3, Polished: 3}}
\label{sec:expe}

\subsection{Setup: \skug{Full: 5, Polished: 4}}
    \label{sec:setup}

@@

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules}
        \label{fig:mem-usages-onlyvalid}
        \vspace{-0.3cm}
    \end{figure}

@@

\section{Conclusion  \skug{Full: 0, Polished: 0}}
\label{sec:conc}
